Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2020-07-06 paper/ifo-appendix.tex
editing wiener filter text from Changrong

\han{to do! pseudo-code here}
\jam{[Hannah has already explained a lot about Viterbi, at this point is it worth just citing existing documentation for further details? I used Wikipedia in writing code, should we cite that? Andrew mentions going through the practical aspects, e.g. log to avoid underflow, is that worth writing up? My thought is only that all of this is technically available in the open source.]}
%We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time. These are normalised between $(0, 1)$ to use as multiplicative weights. In fact, we take the logarithm and so the sum of all weights to avoid underflow. The best path has the highest weight among all paths.

@@

Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram. 
It is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}. A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}. This design was based off \jam{standard examples (I have no idea where I or Alex originally got the designs from, I'm confident though that they are common enough to not need citing given that we also give the diagram)} of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.

>>>>

\han{to do! pseudo-code here}
\jam{[Hannah has already explained a lot about Viterbi, at this point is it worth just citing existing documentation for further details? I used Wikipedia in writing code, should we cite that? Andrew mentions going through the practical aspects, e.g. log to avoid underflow, is that worth writing up? My thought is only that all of this is technically available in the open source.]}
\han{[from Hannah: I think it is worth having a go at pseudocode write-up of the implementation as Andrew suggests - James can you add some text based on your code and we can see how it looks?]}
%We consider a spectrogram grid where each element in the grid is the Fourier amplitude of a particular frequency at a particular time. These are normalised between $(0, 1)$ to use as multiplicative weights. In fact, we take the logarithm and so the sum of all weights to avoid underflow. The best path has the highest weight among all paths.

@@

Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram. 
It is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}. A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}. This design was based off \jam{standard examples (I have no idea where I or Alex originally got the designs from, I'm confident though that they are common enough to not need citing given that we also give the diagram)} \han{[from hannah: hmm, okay - I think if we refer to `standard examples' we need to say what that means. Are there any equivalent references you can find which include the same information, even if you can't find exactly the one you used? If we think about someone who may like to reproduce this work, this is a very useful piece of information]} of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.

### 2020-07-06 paper/ifo-complexAudio.tex
editing wiener filter text from Changrong

We use vector $\textbf{x}=\{x(0),\dots x(N-1)\}$ to denote the noisy speech signal sequence. Assume original clean signal is $\textbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\textbf{w}=\{w(0),\dots,w(N-1)\}$. We then have
\begin{equation}
    \textbf{x}=\textbf{s}+\textbf{w}
\end{equation}
Based on the observation data$\textbf{x}$, our goal is to estimate s, such that the error between $\textbf{x}$ and estimated $\mathit{\hat{s}}$ is minimized in the mean square sense, we write
\begin{equation}
    \hat{\textbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\textbf{x}-\textbf{s}||_2^2
\end{equation}
Under the assumptions that observed data is wide sense stationary and noise sequence is stationary, uncorrelated with signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
The famous Wiener-Hopf equation \citep{noble1959methods} provides us a way to determine $\textbf{h}$, which can be expressed as 
\begin{equation}
\label{winer-hopf}
\begin{bmatrix}  
r_{xx}[0]&r_{xx}[1]&\dots& r_{xx}[n]\\
r_{xx}[1]&r_{xx}[0]&\dots &r_{xx}[n-1]\\
\vdots&\vdots&\ddots&\vdots\\
r_{xx}[n]&r_{xx}[n-1]&\dots &r_{xx}[0]
\end{bmatrix}
\begin{bmatrix}
h[0]\\
h[1]\\
\vdots\\
h[n]
\end{bmatrix}=
\begin{bmatrix}
r_{ss}[0]\\
r_{ss}[1]\\
\vdots\\
r_{ss}[n]
\end{bmatrix}
\end{equation}

\begin{equation*}
\left\{
\begin{array}
r_{xx}[n]=E[x[i]x[i+n]] \\ 
r_{ss}[n]=E[s[i]s[i+n]]
\end{array} \right\}  
\end{equation*}
where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.
If further, we do not need to guarantee the causality of Wiener filter, in other words, we estimate current signal based on both past and future observations, Eq. \ref{winer-hopf} can then be represented in the frequency domain, which is
\begin{equation}
\label{wiener}
    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}
\end{equation}
where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ stands for the frequency spectrum of noisy observations, clean signal and pure noise respectively. Intuitively, from Eq. \ref{wiener}, we can see Wiener filter amplifies input signal where signal to noise ratio (SNR) is high, and weaken the signal where SNR is low.
Notice that since this filter is non-causal, which may not be suitable for real time signal processing. More detailed analysis can be found in \citep{10.5555/151045}. In our case, we construct a causal Wiener filter based on Eq. \ref{winer-hopf}, and set the length of the filter to be 100. Large n smooths the input signal further, while cause more delay and requires more memory meanwhile. $n=100$ is a relatively reasonable choice for balancing these effects.

%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
%\han{should we add more details here too?} \jam{[seconded]}
%The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
%The timeseries is visibly cleaner, however there is a strong noise hum.

>>>>

A Wiener filter is used to compute an estimate of the original signal. 
We label the noisy speech signal sequence as  $\mathbf{x}=\{x(0),\dots, x(N-1)\}$, where $N$ is the length of the data sequence. 
It is the sum of the original clean signal $\mathbf{s}=\{s(0),\dots,s(N-1)\}$, and the noise sequence $\mathbf{w}=\{w(0),\dots,w(N-1)\}$, 
\begin{equation}
    \mathbf{x}=\mathbf{s}+\mathbf{w}
\end{equation}
Given the observation data $\mathbf{x}$, our goal is to estimate $\mathbf{s}$ so that the error between $\mathbf{x}$ and the estimate ${\hat{\mathbf{s}}}$ is minimized in the mean square sense,
\begin{equation}
    \hat{\mathbf{s}}=\mathop{\arg\min}_{\textbf{s}}||\mathbf{x}-\mathbf{s}||_2^2\,. 
\end{equation}
Assuming that $\textbf{x}$ is wide-sense stationary \han{[perhaps a simple definition of wide-sense stationary should be included in brackets here? to do; ask Changrong]} and that $\textbf{w}$ is stationary and uncorrelated with the signal, we represent the $n$th order Wiener filter by its coefficients $\textbf{h}=\{h(0),\dots(n)\}$. 
The famous Wiener-Hopf equation~\citep{noble1959methods} provides us a way to determine $\textbf{h}$, as
\begin{equation}
\label{eqn:winer-hopf}
\begin{bmatrix}  
r_{xx}[0]&r_{xx}[1]&\dots& r_{xx}[n]\\
r_{xx}[1]&r_{xx}[0]&\dots &r_{xx}[n-1]\\
\vdots&\vdots&\ddots&\vdots\\
r_{xx}[n]&r_{xx}[n-1]&\dots &r_{xx}[0]
\end{bmatrix}
\begin{bmatrix}
h[0]\\
h[1]\\
\vdots\\
h[n]
\end{bmatrix}=
\begin{bmatrix}
r_{ss}[0]\\
r_{ss}[1]\\
\vdots\\
r_{ss}[n]
\end{bmatrix}
\end{equation}
where $r_{xx}$ and $r_{ss}$ are the auto-correlation functions of $\mathbf{x}$ and $\mathbf{s}$ between timestep $i$ and $i+n$, 
%\begin{equation*}
%\left\{ 
\begin{eqnarray} 
r_{xx}[n] &~=~& E[x(i)~x(i+n)] \,,\\
r_{ss}[n] &~=~& E[s(i)~s(i+n)] \,.
%r_{xx}[n] &=& E[x(i)\, &x(i+n)] \\ 
%r_{ss}[n] &=& E[s(i)\, &s(i+n)] 
\end{eqnarray} 
%\right\}  
%\end{equation*}
%where $r_{xx}$ and $r_{ss}$ is the auto-correlation function of $x$ and $s$.
If we do not need to guarantee the causality of the Wiener filter, i.e. we can estimate the current signal based on both past and future observations, then Eqn.~\ref{eqn:winer-hopf} can be represented in the frequency domain as
\begin{equation}
\label{eqn:wiener}
    H(f)=\frac{P_{ss}(f)}{P_{xx}(f)}=\frac{P_{ss}(f)}{P_{ss}(f)+P_{ww}(f)}
\end{equation}
where $P_{xx}(f),P_{ss}(f),P_{ww}(f)$ stands for the frequency spectrum of the noisy observation, clean signal, and pure noise respectively. 
Intuitively, from Eqn.~\ref{eqn:wiener}, we can see that the Wiener filter amplifies the input signal where signal to noise ratio (SNR) is high, and weakens the signal where SNR is low.
We note that this filter is non-causal, which may not be suitable for real time signal processing. 
More detailed analysis can be found in Ref.~\citep{10.5555/151045}. 
In this work, we construct a causal Wiener filter based on Eqn.~\ref{eqn:winer-hopf}, and set the length of the filter to $n=100$. 
A larger $n$ provides more smoothing of the input signal, however requires more computational memory. 
The choice of $n=100$ is a reasonably selection to balance smoothing and efficiency.

The timeseries and frequency spectrum results of the Wiener fileter are are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
There is improvement which is particularly visible in the timeseries, however a strong noise hum persists. 
%The Wiener filter produces an estimated signal $\hat{x}$ using linear time-invariant filtering. 
%It minimizes the mean square error $\mathbb{E}[(\hat{x}-x)^2]$ between $\hat{x}$ and the actual signal $x$~\citep{verhoeven2009robust}.
%\han{should we add more details here too?} \jam{[seconded]}
%The timeseries and frequency spectrum results of the Wiener filter are shown in the fourth panel of Fig.~\ref{fig:timeNotchWiener} and the second panel of Fig.~\ref{fig:freqNotchWiener} respectively. 
%The timeseries is visibly cleaner, however there is a strong noise hum.
